---
title: "THE STRINGR PACKAGE"
subtitle: "by Gwen Rino"
author: CSX 415.1
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  revealjs::revealjs_presentation:
    incremental: true
    center: true
---
```{r echo=FALSE}
library(stringr)
```

## STRINGR  
has four families of functions

- CHARACTER MANIPULATION  
- WHITESPACE TOOLS  
- LOCALE SENSITIVE OPERATIONS  
- PATTERN MATCHING FUNCTIONS  

## CHARACTER MANIPULATION

## `str_length()`

`str_length()` returns the length of a string.  

```{r}
str_length("abcde")
```

## `str_sub()`

`str_sub()` accesses individual characters in a string.  

```{r}
vec <- c("a string", "another string")

# Returns the third character in each string
str_sub(vec, 3, 3)

# Returns characters 3-5 in each string
str_sub(vec, 3, 5)

# Count from right using negative numbers
str_sub(vec, -4, -1)
```


## `str_sub()`

You can use `str_sub()` to modify strings.  

```{r}
vec <- c("a string", "another string")

str_sub(vec, 3, 3) <- "X"
vec
```

## `str_dup()`

You can use `str_dup()` to duplicate strings.  

```{r}
print(vec)
str_dup(vec, c(2,3))
```

## WHITESPACE TOOLS

## `str_pad()`

`str_pad()` pads a string to a given length by adding white space.  

```{r}
x <- c("z", "abcedfg")
str_pad(x, 10) # Default padding is on the left
str_pad(x, 10, "right")
str_pad(x, 10, "both")
```

## `str_pad()`

You can pad with characters other than spaces using the `pad` argument.  

```{r}
x <- c("z", "abcedfg")

str_pad(x, 3, pad="X")
```
<small>*Notice that padding to a length < the length of the string does not truncate the string!</small>

## `str_trunc()`
You can truncate a string to a given length (including a 3 character ellipsis) using `str_trunc()`.  
```{r}
str_trunc("Thisstringisquitelong", 13, "right")
```

## `str_trim()`
The opposite of `str_pad()` is `str_trim()`. It trims leading and trailing white space.

```{r}
y <- c("   a", "b   ", "  c  ")
str_trim(y) # Default trims white space from both sides
str_trim(y, "left")
str_trim(y, "right")
```

## `str_wrap()`
You can use `str_wrap()` to wrap a paragraph of text, finding whitespace breaks such that the width of each line is as similar to the given argument as possible.
```{r}
jabberwocky <- "`Twas brillig, and the slithy toves did gyre and gimble 
in the wabe: All mimsy were the borogoves and the mome raths outgrabe."

str_wrap(jabberwocky, width = 40)
cat(str_wrap(jabberwocky, width = 40))
```

## LOCALE SENSITIVE OPERATIONS

***
A few stringr functions are locale-sensitive, which means that they can perform differently to accommodate different languages.  

The default is always English. You can accommodate different languages by setting the `locale` argument to a two letter ISO-639-1 code.

You can see a complete list of available locales by running `stringi::stri_locale_list()`.

## `str_sort()` and `str_order()`

For example, in Lithuanian, y comes between i and j.

```{r}
str_order(letters)
str_order(letters, locale = "lt")
str_sort(letters, locale = "lt")
```

## `str_to_lower()`

Another example: Turkish has two different letters 'i', one with and one without a dot.

```{r}
x <- "I like horses."
str_to_lower(x)
str_to_lower(x, "tr")
```

