---
title: "THE STRINGR PACKAGE"
subtitle: "by Gwen Rino"
author: CSX 415.1
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  revealjs::revealjs_presentation:
    incremental: true
    center: true
---
```{r echo=FALSE}
library(stringr)
```

## STRINGR  
has four families of functions

- CHARACTER MANIPULATION  
- WHITESPACE TOOLS  
- LOCALE SENSITIVE OPERATIONS  
- PATTERN MATCHING FUNCTIONS  

## CHARACTER MANIPULATION

## CHARACTER MANIPULATION

`str_length()` returns the length of a string:
```{r}
str_length("abcde")
```

## CHARACTER MANIPULATION

`str_sub()` accesses individual characters in a string.  

```{r}
vec <- c("a string", "another string")

# Returns the third character in each string in the vector
str_sub(vec, 3, 3)

# Returns characters 3-5 in each string in the vector
str_sub(vec, 3, 5)

# Can count from right using negative numbers
str_sub(vec, -4, -1)
```

## CHARACTER MANIPULATION

You can use `str_sub()` to modify strings:
```{r}
str_sub(vec, 3, 3) <- "X"
vec
```

## CHARACTER MANIPULATION

You can use `str_dup()` to duplicate strings:
```{r}
str_dup(vec, c(2,3))
```

## WHITESPACE TOOLS

